Visually pleasing knot projections

### 2021-09-28 inst/jma_rebuttal.tex

P10, line 17: ``... a vertical line of symmetry." - Better to say that
this figure has $D_5$ symmetry with five mirror lines.

\textcolor{blue}{Done.}

P11, line 26: It seems that all the functionality needed to deal with
links is already present in the software packages mentioned, as long
as one can let the program go around more than one single closed loop.

\textcolor{blue}{I have spent many many hours pondering how to deal
  with links.  As the referee says, much of the required functionality
  is (in principle) already present in the software.  Each component
  would have its own badness, and in addition there would be
  $n(n-1)/2$ inter-component interaction terms measuring features such
  as closeness between non-intersecting strands of different
  components.  Several problems stood out: firstly, the problem of
  inter-component intersection points.  Currently this is a
  (symmetrical) Boolean matrix with rows and columns corresponding to
  strands, and entry $(i,j)$ being whether or not strand $i$
  intersects with strand $j$.  The link generalization would be a
  symmetric matrix of intersection matrices: an object with four
  indices and entry $(i,j,k,l)$ indicating whether strand $j$ of
  component $i$ intersects strand $k$ of component $l$.  I found this
  object to be unwieldy to work with and difficult to manipulate.  The
  second problem was the imposition of symmetry.  Sometimes one might
  desire that some components of a link have a particular kind of
  symmetry and others to have a different kind (or no!) symmetry.  One
  unexpectedly difficult problem was that one might have a subset of
  links that individually possess no symmetry but collectively possess
  mirror-, rotational-, or indeed dihedral- symmetry.
  Thistlethwaite's L4a1 is problematic: one would expect the two
  components to be identical except for a 90-degree rotation, and I
  was unable to implement this in the context of the package. L6a3 is
  similarly difficult.  We would need new badnesses too.  Consider
  L6a4; the smaller component should be not only as circular as
  possible (?), but also one might desire the overall length to be
  smaller than that of the other component.\\ My considered view is
  that dealing with links would be possible in principle but, due to a
  number of non-interesting technical reasons (which are not really
  suitable for inclusion in a scholarly publication), an order of
  magnitude harder than the single-component links considered in the
  submission.  I'm not saying it's impossible, but just very difficult
  and, for me, firmly in the category of ``further work''.  I hope
  that's OK.}

### 2021-09-28 inst/jma_knot.tex
minor tweaks

It is common to present a knot using diagrams such as
Figure~\ref{knot_table}, in which a two-dimensional projection of the
knot is given with broken lines indicating where one strand passes
over another.  Knot diagrams have a long history:
Przytycki~\cite{przytycki1998} discusses an intricate braid used as a
seal sign in around 1700\,BCE; they figure prominently in {\em The
  Book of Kells} and the Lindisfarne gospels, graphic art produced
around 800\,CE~\cite{bain1973}.  Such Celtic artwork makes extensive
use of symmetry~\cite{cromwell2008}, and their visual appeal persists
to the present day~\cite{antonsen2021}.  More recent examples would
include bobbin lace~\cite{irvine2020}, in which many threads are
braided together to form a loose fabric.  The enduring popularity of
such fibre arts attests to the aesthetic of knots and braids in the
modern world.

@@

Symbolic software such as Mathematica and Maple can be used to produce
representations of knots, typically in the form of nonstandard
embeddings of the torus into~$\mathbb{R}^3$; but as discussed below
this approach is suboptimal from the perspective of producing a
visually appealing projection.

Visual appeal of knot projections appears to be important to
contemporary mathematicians.  Recent work by Taalman (writing as
mathgrrl)~\cite{taalman2019}, for example, shows that large amounts of
effort have been expended by serious mathematicians making
representations of knots attractive.  Robert Bosch~\cite{bosch2010},
taking a different approach to knot representations, presents artwork
that utilises high-dimensional computer optimization in a similar way
to that presented here.

No discussion of computer-generated representations of knots would be
complete without the tour de force that is~{\em
  KnotPlot}~\cite{scharein1997,scharein1998} which includes extensive
functionality for visualising knots.  Much of the visualization
considers (nonstandard) embeddings of a torus in $\mathbb{R}^3$; the
configuration is ``relaxed'' in such a way as to produce a smoother
output.  {\em KnotPlot} has two major display modes: any knot may be
displayed in either a ``beads and sticks" mode or in a ``smooth tubes"
mode.  The software does render knot diagrams but the author states
[p77] that ``knots with higher crossing numbers (greater than six or
seven) will not produce good 2D projections when fully relaxed", and
admits that some of the diagrams are ``a bit messy".

@@

The R programming language~\cite{rcore2021}, while usually associated
with statistical analysis and data processing, is also a suitable tool
for producing visually attractive illustrations.  Generative art [that
  is, use of an algorithm, with or without a random component, in a
  creative process] is possible in R~\cite{brunner2021}.  More
specificially to knot theory, the {\tt Rknots}
package\cite{comoglio2011,comoglio2016} presents software for folded
protein structures, and includes 3D renderings of commonly encountered
knotted strands.

>>>>

It is common to present a knot using diagrams such as
Figure~\ref{knot_table}, in which a two-dimensional projection of the
knot is given with broken lines indicating where one strand passes
over another.  Knot diagrams have a long history:
Przytycki~\cite{przytycki1998} discusses an intricate braid drawn
around 1700\,BC; they figure prominently in {\em The Book of Kells}
and the Lindisfarne gospels, graphic art produced around
800\,AD~\cite{bain1973}.  Such artwork makes extensive use of
symmetry~\cite{cromwell2008}, and its visual appeal persists to the
present day~\cite{antonsen2021}.  More recent examples would include
bobbin lace~\cite{irvine2020}, in which many threads are braided
together to form a loose fabric.  The enduring popularity of such
fibre arts attests to the aesthetic of knots and braids in the modern
world.

@@

Symbolic software such as Mathematica and Maple can be used to produce
representations of knots, typically in the form of nonstandard
embeddings of the torus into~$\mathbb{R}^3$; but as discussed below
this approach is suboptimal from the perspective of producing a
visually pleasing projection.

Visual appeal of knot projections appears to be important to
contemporary mathematicians.  Recent work by Taalman (writing as
mathgrrl)~\cite{taalman2019}, for example, shows that large amounts of
effort have been expended by serious mathematicians making
representations of knots attractive.  Bosch~\cite{bosch2010}, taking a
different approach to knot representations, presents artwork that
utilises high-dimensional computer optimization, as here.

No discussion of computer-generated representations of knots would be
complete without mentioning the tour de force that is~{\em
  KnotPlot}~\cite{scharein1997,scharein1998} which includes extensive
functionality for visualising knots.  Much of the visualization
considers (nonstandard) embeddings of a torus in $\mathbb{R}^3$; the
configuration is ``relaxed'' in such a way as to produce a smoother
output.  {\em KnotPlot} has two major display modes: any knot may be
displayed in either a ``beads and sticks" mode or in a ``smooth tubes"
mode.  The software does render knot diagrams but the author states
[p77] that ``knots with higher crossing numbers (greater than six or
seven) will not produce good 2D projections when fully relaxed", and
admits that some of the diagrams are ``a bit messy".

@@

The R programming language~\cite{rcore2021}, while usually associated
with statistical analysis and data processing, is also a suitable tool
for producing visually attractive images.  Generative art [that is,
  use of an algorithm, with or without a random component, in a
  creative process] is possible in R~\cite{brunner2021}.  More
specificially to knot theory, the {\tt Rknots}
package\cite{comoglio2011,comoglio2016} presents software for folded
protein structures, and includes 3D renderings of commonly encountered
knotted strands.
